---
title: "Goodman Project 1"
output: html_document
---

In my project, I work with mouse adenovirus type 1 (MAV-1). Specifically, I'm looking at a cellular kinase and its potential role in antiviral defense. Canonically, protein kinase R (PKR) responds to the presence of double-stranded RNA and it becomes activated to phosphorylate the transcription factor eIF2alpha. When phosphorylated, eIF2alpha halts protein synthesis. GCN2 (general control nonderepressible 2) is another kinase that phosphorylates eIF2alpha; it's activated by amino acid deprivation and UV irradiation. However, recent studies have shown that GCN2 is also activated by Sindbis virus (Berlanga et al., 2006) and HIV-1 (Cosnefroy et al., 2013). I hypothesize that GCN2 is also important during MAV-1 infection.

To test my hypothesis, I have been working with GCN2 mutant mice, termed Atchoum mice. These mice were created during a screen done by the Beutler group. They used ENU mutagenesis to create a library of C57BL/6-background mice, each with single point mutation. They then harvested peritoneal macrophages from each strain of mouse and screened them for susceptibility to human adenovirus and mouse cytomegalovirus. They found that the Atchoum strain of mice showed increased susceptibility to infection by both human adenovirus and mouse cytomegalovirus. When they sequenced the genome of the Atchoum mouse, they found that the single point mutation was in the GCN2 gene. Further investigation determined that the mutation caused a missense mutation and caused incorrect posttranslational splicing that resulted in an inactive kinase.

To test my hypothesis, I started with attempting to recapitulate the macrophage experiments as done by the Beutler group. I harvested peritoneal macrophages from Atchoum and C57BL/6 mice and plated them in 6 well plates. I infected them the next day with MAV-1 at an MOI of 1. After 24, 48, and 72 hours, I harvested individual wells of cells (each is one replicate) and separated the cells from the supernatant. I purified DNA from both the cell and supernatant fraction and used qPCR to quantify the viral loads.

I get the output in an excel document, which I converted to text documents, one for the cell fraction data and one for the supernatant fraction data. The first six lines of each are listed below, cell data first. 

```{r echo=FALSE}
#load data into r studio
celldata <- read.table(file="qPCR_Raw_Data_Cells.txt", header=T)
superdata <- read.table(file="qPCR_Raw_Data_Supernatant.txt", header=T)
head(celldata)
head(superdata)
```

Let's start with the cell fraction data. I don't get quantifiable levels of DNA from the cell fraction so I have to use housekeeping genes to normalize the amount of DNA over all of the samples. I use GAPDH, so I get GAPDH output in my data. Since I use the qPCR machine program to normalize the data, I no longer need the GAPDH information, I only need the mE1A data (viral load data). So I want to delete all of the rows that contain GAPDH data. The first six rows of the modified data set are shown below.

```{r echo=FALSE}
#remove GAPDH data
celldata <- celldata[celldata$target_name == "mE1A",]
head(celldata)
```

Now I can focus on the viral load data. I'm going to separate the Atc and B6 data into two different tables, using the same method I did to remove the GAPDH data. See the first six lines of each of the separated data below.

```{r echo=FALSE}
atc_celldata <- celldata[celldata$strain == "Atc",]
b6_celldata <- celldata[celldata$strain == "B6",]
head(atc_celldata)
head(b6_celldata)
```

When I load my samples into the qPCR plate, I do three replicates for each sample. So in the data output, there are three rows of data for each sample. I want to average the RQ data because that is the relevant data (it is the fold change in viral load with the C57BL/6 data set as a basepoint). The first six rows of the averaged B6 data set are shown below.

```{r echo=FALSE}
#average RQ data
avg_atc_celldata <- aggregate(.~sample_name, FUN=mean, data=atc_celldata)
avg_b6_celldata <- aggregate(.~sample_name, FUN=mean, data=b6_celldata)
head(avg_b6_celldata)
```

After averaging the RQ data, I need to aggregate the data by time point. 

```{r echo=FALSE}
#average the data by time point
group_atc_celldata <- aggregate(.~time_point, FUN=mean, data=avg_atc_celldata)
group_b6_celldata <- aggregate(.~time_point, FUN=mean, data=avg_b6_celldata)
```

Now that I have the average of all my samples for each of the time points, I want to plot the data so I can see it.

```{r echo=FALSE}
#plot the data
par(mfrow = c(1,1))
barplot(group_atc_celldata$RQ, names.arg=group_atc_celldata$time_point, xlab="Hours Post Infection", ylab="Fold Change Over C57BL/6")
barplot(group_b6_celldata$RQ, names.arg=group_b6_celldata$time_point, xlab="Hours Post Infection", ylab="Fold Change Over C57BL/6", add=T)
```

Now that I've visualized the data, I have an idea of whether the difference between the viral loads in the Atchoum versus C57BL/6 macrophages is going to be significantly different. I need to do some t-tests to confirm my hypothesis. I will do two sided t-tests for each time point, comparing the Atchoum macrophage viral loads to the C57BL/6 macrophage viral loads. See t-test outputs below, in order from 24 hours post infection to 72 hours post infection.

```{r echo=FALSE}
#analyze data
t.test(avg_atc_celldata$RQ[avg_atc_celldata$time_point == 24], avg_b6_celldata$RQ[avg_b6_celldata$time_point == 24])
t.test(avg_atc_celldata$RQ[avg_atc_celldata$time_point == 48], avg_b6_celldata$RQ[avg_b6_celldata$time_point == 48])
t.test(avg_atc_celldata$RQ[avg_atc_celldata$time_point == 72], avg_b6_celldata$RQ[avg_b6_celldata$time_point == 72])
```

That completes my analysis of the cellular fraction. Now I want to look at the supernatant data. I'm going to start again by separating the Atchoum and C57BL/6 data. See the head of the C57BL/6 table below.

```{r echo=FALSE}
atc_superdata <- superdata[celldata$strain == "Atc",]
b6_superdata <- superdata[celldata$strain == "B6",]
head(b6_superdata)
```

Again, like with the cellular data, I do three replicates of each sample in the qPCR plate so I need to average each of these replicates to get a single viral load quantity for each sample. Since there is no cellular DNA in this fraction, I did not have to worry about having housekeeping genes for normalization. I just purified DNA from the same amount of supernatant for each sample and I plated the same amount of purified DNA for each sample and just quantitated the viral load directly. See the head of the averaged C57BL/6 data table below.

```{r echo=FALSE}
#average the data replicates quantity
avg_atc_superdata <- aggregate(.~sample_name, FUN=mean, data=atc_superdata)
avg_b6_superdata <- aggregate(.~sample_name, FUN=mean, data=b6_superdata)
head(avg_b6_superdata)
```

```{r echo=FALSE}
#average the data by time point
group_atc_superdata <- aggregate(.~time_point, FUN=mean, data=avg_atc_superdata)
group_b6_superdata <- aggregate(.~time_point, FUN=mean, data=avg_b6_superdata)
```

